---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(readr)
```

# Lab report #3 - Instructions

Follow the instructions posted at <https://ds202-at-isu.github.io/labs.html> for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.

All submissions to the github repo will be automatically uploaded for grading once the due date is passed. Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`.

```{r}
# Deaths per avenger
deaths <- av |> 
  select(Name.Alias, starts_with("Death")) |> 
  pivot_longer(
    Death1:Death5,
    names_to = "Time",
    values_to = "Death"
  ) |>
  mutate(
    Time = parse_number(Time)
  ) |>
  filter(Death != "")

head(deaths)
```


Similarly, deal with the returns of characters.
```{r}
# Returns per avenger
returns <- av |>
  select(Name.Alias, starts_with("Return")) |> 
  pivot_longer(
    Return1:Return5,
    names_to = "Time",
    values_to = "Return"
  ) |>
  mutate(
    Time = parse_number(Time)
  ) |>
  filter(Return != "")

head(returns)
```

Based on these data sets calculate the average number of deaths an Avenger suffers.
```{r aver_deaths}
# Average number of deaths per Avenger
avg_deaths <- deaths |> 
  filter(Death == "YES") |>  
  group_by(Name.Alias) |>
  summarise(death_count = n()) |>
  summarise(avg_deaths = mean(death_count))

avg_deaths
```
The average number of deaths an Avenger suffer's is 1.39

## Individually

For each team member, copy this part of the report.

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.




## Betsy's Work

Quoted statement I plan to fact-check:
> "Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team"

Code to derive the (numeric) fact for the statement:
```{r}
deaths |> 
  group_by(Name.Alias) |>
  summarise(death_count = n()) |>
  filter(death_count == 1) |> 
  summarise(one_death = n())
```

The result:
According to the dataset, there are 145 Avengers with exactly one death according to this data set which differs from FiveThirtyEight's analysis of 69 Avengers with one death.


## Maurycy's Work

Quoted statement I plan to fact-check:
> "I counted 89 total deaths — some unlucky Avengers are basically Meat Loaf with an E-ZPass — and on 57 occasions the individual made a comeback."

Code to derive the (numeric) fact for the statement:
```{r}
# Total deaths
total_deaths <- deaths |> 
  filter(Death == "YES") |> 
  nrow()

total_deaths
```

```{r}
# Total returns
total_returns <- returns |> 
  filter(Return == "YES") |> 
  nrow()

total_returns
```

The result:
According to the dataset, there are 89 total recorded deaths and 57 returns, which matches the numbers reported in the FiveThirtyEight analysis.


## Anthony's Work

Quoted statement I plan to fact-check:
> "The MVP ... has to be Jocasta ... who has been destroyed five times and then recovered five times."

```{r}
# Deaths for Jocasta
jocasta_deaths <- deaths |>
  filter(Death == "YES", Name.Alias == "Jocasta") |>
  summarise(total_deaths = n())

# Returns for Jocasta
jocasta_returns <- returns |>
  filter(Return == "YES", Name.Alias == "Jocasta") |>
  summarise(total_returns = n())

list(
  deaths = jocasta_deaths$total_deaths,
  returns = jocasta_returns$total_returns
)
```
The result:
According to the data set, there are 5 deaths and 5 returns for Jocasta, so the statement is indeed true!
